---
title: "esm_206_lab_5"
author: "Garrett Eyer"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Learn to parse dates with 'lubridate'
- Get counts of observations with count()
- Uncount() rows 
- One and two-sample t-test
- Create a heatmap with geom_tile()

Attach 

```{r, include = FALSE}
# Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here("data", "lobster_abundance.csv"), 
                              na = "-99999") %>% 
  clean_names()

# Data is shown in frequency format, needs to be in case format...shows counts for specific area and groups them...want to ungroup so we can measure data easier - we want every observation to have its own row
```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format.

```{r}

lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```

#### Exploratory data visualization 

We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.3,
              width = 0.2)


# Histogram 

ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

# Quantile - Quantile Plot...need to use in aes(the sample is THIS)

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq(color = "black",
          alpha = 0.5) +
  facet_wrap(~site) +
  theme_minimal()
  


```


Do these look normally distributed? Yes
Does it make sense that lobster size is normally distributed? Yes

#### Convert the 'date' column to class 'Date'

We'll use the 'lubridate' package to convert to Data format, and then to help us easily parse month & year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )

```

Now let's parse year and month using lubridate::month() and lubridate::year():

```{r}

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )

# label = TRUE, changes month number to the month abbreviation 
# turned date into factor and can see the levels are Aug-July
```


Now let's find counts of observed lobsters based on different hierarchical groupings:

First: let's count lobsters by year and month:
```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)


lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_site_total <- lobster_parse_date %>% 
  dplyr::count(site)


# give it the variables you want to group by then count...for count()
```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by() + n()

```{r}

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )


# group by site then it summarizes based on the site for these variables...thena lso includes the population size with = n 
# Count assumes you want to do group_by and summarize all together
```

#### Find confidence intervals

Use t.test() function to find confidence intervals (for one sample), and perform t-tests to compare means of two samples (...this will be covered conceptually in lecures Week 6)

```{r}

ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm) # pulls it to a single vector, does not keep in dataframe format

t.test(ivee_lobster)

```



#### Two-sample t-test to compare means 

We want to ask: is there a significant difference in lobster lengths at Naples and Mohawk reefs?

We've done our necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate. Comparing means between sites.  Have to give t-tests the entire vector for a sample group...can't just give one item..

In either direction are these significantly different?  That's what a 2 sample t test is asking...

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)

mn_ttest

# For confidence interval..."In 95% of those cases, the true mean will between these ranges..."
# Difference? Why? Prob drawn from populations with different means 
# Low p value means it's not likely we got these different means from the same population 

```


There is a significant difference in lobster lengths betwen Naples and Mohawk Reef.  (t(df) = statistic, p < .001, alpha = .05).


```{r}

lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK")) 

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2$p.value
mn_ttest2$statistic
#this is better than writing it out - inlclude these in Rmd - use in line referencing like below  


```

There is a sig diff (t(1850.8) = -19.849, p = 2.2e-16).  

Don't use this ^ not reproducible...use in-line code to write it

Here is the t-statistic: `r mn_ttest2$p.value`

Use the backticks then r to indicate it's from an r chunk, then put the name where it is stored, then what you're looking for 


```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

